---
title: "ENV S 193DS: Coding Work"
format: html
toc: FALSE 
---

## Overview:

The following code chunks are from a homework assignment for my EVN S 193DS class where our objective was to visualize personal observed data that we'd been collecting throughout the quarter. In the code check below, I'm reading in my necessary packages that I used to organize and work on the assignment. Additionally, I read in my personal data, titled "DataSheet(1) copy.csv)", and turned it into an object titled "ActiveData1" to make it easier to work with.

```{r, message=FALSE}
show_col_types = FALSE

library(tidyverse)
library(here)
library(gt)
library(janitor)

ActiveData1 <- read_csv("DataSheet(1) copy.csv")
```

## Data Wrangling and Calculations

The next code chunk contains the code necessary to wrangle the specific data columns that I want to use for my analysis. I created another object and selected the 'Long Day' and 'Money Spent' columns and calculated the average money spent on long days. I then created another object titled 'average_money_summary' which I used to create a group using the 'Long Day' data. I then summarized calculations of mean money spent, standard deviation, standard error, and the upper and lower quartiles. Lastly, I rounded the values to one decimal place.

```{r}
# creating `average_money_summary` before so R can read it in order to render 
money_spent_long_day <- ActiveData1 |> # creating new object from 'ActiveData1' to 'money_spent_long_day'
  select(`Long Day`, `Money Spent`) # selecting `Long Day` and `Money Spent` 

average_money_summary <- money_spent_long_day |> # create new object from `money_spent_long_day` to `average_money_summary` 
  group_by(`Long Day`) |>  # grouping by `Long Day` points 
  summarize(mean = mean(`Money Spent`), # summarizing the mean from `Money Spent` 
            sd = sd(`Money Spent`), # calculating the standard deviation from `Money Spent` 
            se = sd / sqrt(n()), # calculating standard error from standard deviation 
            ci_lower = mean - qt(0.975, df = n() - 1) * se, # calculating the upper and lower quartile 
            ci_upper = mean + qt(0.975, df = n() - 1) * se) |> 
  
  
  mutate(across(c(mean, sd, se, ci_lower, ci_upper), # round mean, sd, se, ci upper and lower to one decimal place 
                ~ round(.x, 1)))
```

## Creating a Table

The following code chunk expresses the information that was calculated above but in a table format where I changed the labels to reflect the information displayed.

```{r}
#| message: false
average_money_summary |> # start with average_money_summary
  gt() |> # calling in the 'gt' table function 
  cols_label( # labeling the different aspects of the table by their full names
    `Long Day` = "Long Day",
    mean = "Mean",
    sd = "Standard Deviation",
    se = "Standard Error",
    ci_lower = "95% Confidence Level (Lower)",
    ci_upper = "95% Confidence Level (Upper)"
    
  ) |> 
  tab_header(title = "Money Spent Summary Statistics") # setting the title as "Money Spent Summary Statistics" 


```

## Visualizing the Data

The following code chunk contains the code needed to make a visualization of my personal data along with the calculations that I made earlier (standard deviation, standard error, upper and lower quartiles). I started by using the data object that I created earlier to create underlying data set while placing my original data (ActiveData1) over it. I also added an error bar displaying the standard error and mean.

```{r}
#| message: false

ggplot(data = average_money_summary, # starting with data frame 
       aes(x = `Long Day`, # label x-axis as 'Long Day'
           y = `mean`, # label y-axis as 'mean' 
           fill = `Long Day`)) + # fill color for 'Long Day' variable 
  geom_jitter(data = ActiveData1, # use jitterplot function starting with 'ActiveData1' data frame 
              aes(x = `Long Day`, # set x-axis as Long Day for jitterplot 
                  y = `Money Spent`, # set y-axis as Money Spent  
                  color = `Long Day`), # use color for Long Day variable
              shape = 18, # setting open circle for shape 
              alpha = 1, # setting transparency 
              size = 2, # setting size of points 
              height = 0, # disabling vertical jitter points 
              width = 0.1) + # enabling horizontal jitter points 
  geom_errorbar(aes(ymin = mean - se, # creating errorbar and creating the upper and lower edges of the standard error 
                    ymax = mean + se),
                width = 0.1, # setting width of the caps
                color = "black" # setting color to black 
                ) + 
  geom_point(data = average_money_summary, # adding primary data frame to display on top of the previously displayed data 
             aes(x = `Long Day`, # setting x-axis as 'Long Day'
                 y = `mean`), # setting y-axis as mean money spent   
             inherit.aes = FALSE, # seperating new aesthetics from the previous aesthetics 
             color = "black", # setting color to black for mean point 
             size = 2) + # setting size to 2 
  theme(panel.background = element_rect(fill = "white"), # filling background with white to disable panels 
        panel.grid.major = element_line(color = NA), # disabling major panel lines 
        panel.grid.minor = element_line(color = NA), # disabling minor panel lines 
        legend.position = "none", # disabling the legend from showing up 
        axis.line = element_line(color = "black")) + # setting color of axis lines to black 
  scale_color_manual(values = c("Yes" = "dodgerblue", # setting color of 'yes' money points to dodgerblue 
                                "No" = "firebrick")) + # setting color of "no" money points to firebrick 
          
  # manually setting title to 'Money Spent on Campus with Standard Deviation and Standard Error', adding subtitle to 'By Mark Oregel', changing y-axis name to 'Money Spent (dollars)', changing x-axis name to 'Long Day'. 
  labs(title = "Money Spent on Campus with Standard Deviation and Standard Error",
       subtitle = "By Mark Oregel",
       y = "Money Spent (dollars)",
       x = "Long Day") 
  

```

### Caption:

This jitter plot displays a comparison of money spent on campus between long and short days. Each point represents a daily observation with the red and blue dots indicating individual spending. The black points represent the group means and the error bars show the standard error of the mean.
